An Elastica-driven Digital Curve Evolution Model for Image Segmentation\thanks{This work has been partially funded by CoMeDiC ANR-15-CE40-0006 research grant.}

### 2019-07-07 jmiv.tex
Suggestions acceptance and corrections

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage[utf8]{inputenc}
\usepackage{graphicx,subfig}
\usepackage{amstext,amsmath,amssymb,bm,bbm,mathtools}
\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
\usepackage[export]{adjustbox}
\usepackage{xcolor}

@@

\newcommand{\todo}[1]{{\textcolor{blue}{#1}}}
\newcommand{\test}[1]{{\textcolor{red}{#1}}}
\newcommand{\jaco}[1]{{\textcolor{green!50!black}{#1}}}
\newcommand{\HT}[1]{{\textcolor{violet}{#1}}}

@@

\begin{document}
%
\title{A\HT{n Elastica-driven} Digital Curve Evolution Model for Image Segmentation\thanks{This  work has  been  partially  funded by CoMeDiC ANR-15-CE40-0006 research grant.}}
  \titlerunning{Digital Curve Evolution by Elastica Energy}
  %\title{Digital Curvature Evolution Models for Image Segmentation\thanks{This  work has  been  partially  funded by CoMeDiC ANR-15-CE40-0006 research grant.}}

\author{Daniel Antunes\inst{1}
Jacques-Olivier Lachaud\inst{1}
Hugues Talbot\inst{2}}
%
\authorrunning{D. Antunes et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Universit{\'e} Savoie Mont Blanc, LAMA, UMR CNRS 5127, F-73376, France
\email{daniel.martins-antunes, jacques-olivier.lachaud@univ-savoie.fr} \and
CentraleSupelec Inria \'equipe OPIS, Universit\'e Paris-Saclay, 9 rue Joliot-Curie, F-91190 Gif-sur-Yvette \\
\email{hugues.talbot@centralesupelec.fr}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  \HT{Geometric priors have been shown to be useful in image segmentation to regularize results. For example, the classical
    Mumford-Shah functional uses regions perimeter as prior. This has inspired much research in the last few decades,
    with classical approaches like the Rudin-Osher-Fatemi and most graph-cut formulations, which all use a weighted or
    binary perimeter prior. It has been observed that this prior is not suitable in many applications, for example for
    segmenting thin objects or some textures, which may have high perimeter/surface ratio. Mumford observed that an
    interesting prior for natural objects is the Euler Elastical model, which involves the squared curvature. In other
    areas of science, researchers have noticed that some physical binarization processes, like emulsion unmixing can be
    well-approximated by curvature-related flow like the Willmore flow. 
    However, curvature-related flows are not easy to compute because curvature is difficult to estimate accurately, and the
    underlying optimisation processes are not convex. In this article, we propose to formulate a digital flow that
    approximates a curvature-related flow using a multigrid-convergent curvature estimator, within a discrete
    variational framework. We also present an application of this model as a post-processing step in a segmentation
    framework.}
\keywords{Multigrid convergence  \and Digital estimator \and Curvature \and Shape Optimization \and Image Segmentation.}
\end{abstract}
%
%
%
\setcounter{footnote}{0}
\section{Introduction} %% \todo{(To be extended)}}

Geometric quantities are particularly useful as regularizers in
low-level image analysis, especially when little prior information is
known about the shape of interest. Length penalization is a \HT{well-behaved,} general
purpose regularizer and \HT{many} models \HT{in the literature} make use
of it, from active contours \cite{casseles97geodesic} to level-set
formulations \cite{malladi1995image,malladi1995shape}. Discrete
graph-based variational models have been particularly succesfull to
incorporate length penalization as a penalizer, while keeping the
ability to extract \HT{a} global optimum
\cite{boykov01graphcut,appleton05geodesic}.

However length regularization shows limitations when segmenting
small or thin and elongated objects, as it tends to shrink solutions
or \HT{yields} disconnected solutions. Such drawbacks can be problematic in
image segmentation or image inpainting. It is thus rather natural to
consider curvature (especially squared curvature) \HT{as} a potential
regularizer. Energies involving \HT{both length and} squared curvature are often called
{\em Euler elastica} model (\HT{that} dates back to Euler) or {\em Willmore energy}
($n$-dimensional extension) and have been brought back to attention in
computer vision by Mumford \cite{mumford1994elastica}. A similar
regularizer is also present in the second-order regularizer
of the \HT{original} snake model \cite{kass1988snakes}. \HT{Similarly,
an explicit curvature term was often employed in early level-set methods,
like~\cite{malladi1995image,malladi1995shape,ballester01filljoint} for segmentation or inpainting,}
\HT{However, Casselles {\em et al.} in~\cite{casseles97geodesic} observed that these
terms were not geometric, in the sense that they depended on the discretization}.

@@

One of the first successful uses of \HT{geometric squared} curvature in image processing is the inpainting algorithm
described in \cite{masnou98inpainting}\HT{, where authors} evaluate the elastica energy along the level lines of a
simply-connected image to reconstruct its occluded parts. The non-intersection property of level lines \HT{induces} the
construction of an efficient dynamic programming algorithm. \HT{However, } the curvature energy is \HT{only} coarsely
approximated.  \cite{chan02elasticainpainting} \HT{is} another geometric inpainting method involving elastica
\HT{optimized with level-sets}.

In image segmentation, Zehiry and Grady have shown that injecting curvature as a regularizer \HT{can help} recover thin
and elongated objects \cite{zehiry10fast}. Similarly Schoenemann {\em et al.}  \cite{schoenemann09linear} have
considered the elastica energy in their ratio-based image segmentation approach. In~\cite{schoenemann2011elastic},
Schoenemann~{\em et al.} extend ratio-cut approaches to segmentation with a curvature term. Their global optimization
framework shows the considerable \HT{advantages} of using such regularizer in common binary image segmentation tasks.
However the time complexity of their algorithm \HT{is prohibitive,} even if curvature is \HT{again} coarsely
approximated.  In more recent work, Lim {\em et al.} propose an edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set formulation.  They observe improved results compared
to~\cite{casseles97geodesic} and other perimeter-based priors, but observe that a shape prior and a close-enough
solution must also be provided to achieve good results.

In fact, it is still a very challenging task to inject curvature in
the context of image segmentation.  
\HT{State-of-art methods are in practice} difficult to optimize and \HT{do not scale}
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In
order to achieve reasonable running times, such approaches make use of
coarse curvature estimations for which the approximation error is
\HT{often} unknown. Improving the quality of the curvature estimator has an
important impact on the accuracy of the results, but is
computationally too costly in these methods.

Segmentation methods using Elastica or Willmore energies \HT{generally
reduce to} computing the corresponding flow, i.e. following its
gradient. Well-founded level-set methods involving Willmore energy
\cite{droske2004level} have been used in medical image segmentation
\cite{lim2012introducing}. Another level-set formulation similar to
the Chan-Vese model \HT{was} proposed in \cite{zhu2013image}. Willmore
flow can also be carried out by phase-field models (see
\cite{bretin2015phase}) but they are less suited to image segmentation
since interfaces are blurred in such models. Threshold dynamics can
also be considered for such energies \cite{esedoglu2008threshold} and
has been proposed for image desocclusion \cite{esedoglu2005threshold}.

In this paper we are interested in purely discrete variational segmentation models involving an elastica
energy. Recently, new estimators of curvature along digital contours have been proposed
\cite{roussillon11mdca,coeurjolly13integral,schindele17mdca} with the \HT{desirable} multigrid convergence property. This
motivates us to \HT{propose} models in which they can be used successfully.

\HT{In the remainder, we} investigate the use of the digital integral invariant
curvature estimator \cite{coeurjolly13integral} in a discrete
variational segmentation framework. More precisely, we show how to
incorporate it in a digital flow minimizing its squared
curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO
algorithm \cite{rother07qpbo}. \HT{Our solution does not} compute a global
solution to the whole domain, but our approach leads to a digital flow
similar to continuous \HT{Elastica}, as shown by our experiments. We
present an application of the model as a post-processing step in a
segmentation framework, and demonstrate how it can improve standard
results. The code is freely available on
github\footnote{https://www.github.com/danoan/BTools}.

%% In this work, we investigate the use of a more suitable curvature
%% estimator with multigrid convergent property and its application as a
%% boundary regularizer in a digital flow minimizing its squared
%% curvature. Our method decreases the elastica energy of the contour and
%% its evolution is evaluated on several digital flows.

%% Finally, we
%% present an application of the model as a post-processing step in a
%% segmentation framework. The code is freely available on
%% github\footnote{https://www.github.com/danoan/BTools}.

\begin{itemize}
\item{Geometric properties and regularization}
\item{The role of curvature}

%  Quote \cite{schoenemann2011elastic} for the interest of using curvature energy when segmenting regions in gradient-based data fitting term. It can look for global optimum within a graph representation of the image. Main problem of their approach is the computational complexity as well as rather coarse approximations of curvature. (Be more precise here).

\item{What are the difficulties in use curvatue as a regularizer (point out convex relaxation approaches) and how a digital approach may help us }
\item{Outline: We present two models that illustrate how digital estimators can be used to minimize the elastica energy on digital sets.  }
\item{\todo{Add related works : geometric priors, binary segmentation, other digital curve evolution models}}
\item{\todo{Talk about Willmore flow with phase field or with threshold dynamics}}

%%\item{\textbf{Classical work on elastica and Wilmore flow in imaging}}
  %%\begin{itemize}
  %% \item{Level lines based disocclusion \cite{masnou98inpainting} }	%% DONE
 %%  \item{Filling-In by Joint Interpolation of Vector Fields and Gray Levels \cite{ballester01filljoint} } %% DONE
 %%  \item{Eulers Elastica and Curvature Based Inpaintings \cite{chan02elasticainpainting} } %% DONE
 %% \item{Discrete willmore flow \cite{bobenko2005discrete}}
 %%  \item{Introducing Willmore flow into level set segmentation of spinal vertebrae \cite{lim2012introducing}}	 %%% DONE
 %% \end{itemize}

%%\item{\textbf{Level-sets}}
 %% \begin{itemize}
  %% \item{Image processing via level set curvature flow \cite{malladi1995image} } %% DONE
  %% \item{Shape Modeling with Front Propagation: A Level Set Approach \cite{malladi1995shape} }	%% DONE
%%	\end{itemize}

\item{\textbf{Convex relaxation}}
  \begin{itemize}
  \item{A convex lower semicontinuous approximation of Eulers elastica energy \cite{bredies15convex}}
  \item{Introducing Total Curvature for Image Processing \cite{goldluecke11totalcurvature}}
  \end{itemize}

\item{\textbf{Constraint optimization}}
  \begin{itemize}
  \item{Curvature regularity for region-based image segmentation and inpainting: A linear programming relaxation \cite{schoenemann09linear} }	
  \item{A Fast Algorithm for Euler's Elastica Model Using Augmented Lagrangian Method \cite{tai11elastica} }		
  %%\item{The elastic ratio: Introducing curvature into ratio-based image segmentation \cite{schoenemann2011elastic}
  %%}		%% DONE
  \item{Curvature Regularization for Curves and Surfaces in a Global Optimization Framework \cite{strandmark11globalframework} }
  \end{itemize}

\item{\textbf{Discrete approaches}}
  \begin{itemize}
  \item{Fast global optimization of curvature \cite{zehiry10fast}	}
  \item{Efficient Squared Curvature \cite{nieuwenhuis14efficient} 	}
  \item{Contrast driven {E}lastica for image segmentation \cite{el2016contrast} }	
  \end{itemize}

\end{itemize}

@@

\section{Multigrid convergent estimators \todo{(Mostly copy/paste)}}

The goal of this section is to  introduce the concept of multigrid convergenc and its potential when analyzing digital images with variational models involving geometric quantities. More precisely we wish to delineate shapes within digital images with some priors related to their unknown continuous geometry. Of course we only observe shape digitizations within images.
%% \begin{itemize}
%% 	\item{Why to use multigrid convergent estimators}
%% 	\item{Multigrid convergence for local estimators}
%% 	\item{Reference to examples of multigrid digital estimators(for curvature, length, tangent)}
%% 	\item{Define the elastica energy estimator}
%% 	\item{\test{(Jaco) see if two opposite II evaluations give convergent curvature}}
%% \end{itemize}

%% \hrulefill

@@

\begin{figure}
\center
	\subfloat[\label{}]{%
	\includegraphics[scale=0.2]{images/local_search/definitions/model-regions-square.eps}
	}\hspace{40pt}%
	\subfloat[\label{}]{%
	\includegraphics[scale=0.2]{images/local_search/definitions/glued-curve.eps}
	}%	
	\caption{The blue pixels in figure (a) illustrates a $2$-ring set, while the green pixels its inner-pixel boundary. In figure (b) we highlight an element of set $\mathcal{G}_{11}(C_1,C_2)$. The curve segments of $C_1$ and $C_2$ are colored in red and green, while the junction linels in blue.}
\end{figure}

@@

\begin{figure}[!ht]
\center
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h1/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h05/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h1/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h05/summary-flow.eps}
	}%
		\caption{Local combinatorial optimization process results for the square and flower shapes. Shapes displayed at every $2$ iterations for the square and $7$ iterations for the flower.}	
		\label{fig:local-comb-square-results}
\end{figure}

@@

\begin{figure}
\begin{minipage}[b]{0.5\textwidth}
\center
\includegraphics[scale=0.35]{images/local_search/square-flower.eps}
\caption{Elastica computation along iterations of algorithm \ref{alg:local-search}}
\label{fig:plot-elastica-local-search}
\end{minipage}\hspace{40pt}%
\begin{minipage}[b]{0.45\textwidth}
\center
\captionsetup{type=table}
\begin{tabular}{|l|c|c|}
\hline
& $h=1.0$ & $h=0.5$\\
\hline
Square & 47s (5s/it) & 260s (12s/it)\\
\hline
Flower & 1235s (23s/it)  & 8516s (52s/it)\\
\hline
\end{tabular}
\caption{Running times for the local combinatorial optimization algorithm with $k=1,N=20$. Four threads were used.}
\label{tab:summary-local-comb-rtime}
\end{minipage}
\end{figure}

@@

We define $O_{r}^{(i)}(y) := O^{(i)} \cap B_r(y)$ and an analogous definition holds for $F_{r}^{(i)}(y)$. Expanding \eqref{eq:curvature-estimator-pixels}, we get 
\begin{align*}
  \hat{\kappa}^2(y) &= c_1\Big( c_2 - \sum_{x_j \in B_r(y)} \big( {x_j} + |F_{r}^{(i)}(y)| \big) \Big)^2 \\
   &= c_1 \cdot \Big( C + 2\left( |F_{r}^{(i)}(y)| - c_2 \right) \hspace{-2mm}\sum_{x_j \in O_{r}^{(i)}(y)}\hspace{-2mm}{x_j} + \hspace{-2mm}\sum_{x_j \in O_{r}^{(i)}(y)}\hspace{-2mm}{x_j^2} + \hspace{-2mm}\sum_{ \substack{x_j,x_k \in O_{r}^{(i)}(p) \\ j<k} }\hspace{-2mm}{2x_jx_k}  \Big),
\end{align*}
where $C=c_2^2 - 2c_2 \cdot |F_{r}^{(i)}(y)| + |F_{r}^{(i)}(y)|^2$ is a constant. In
order to be symmetric, the curvature at some linel $y$ is computed on
$R_m(S)$.\todo{Que veux-tu dire?} By ignoring constants and multiplication factors and using
the binary character of the variables, we define the following family
of energies for given parameters $\alpha,\beta \geq 0$.
\begin{align}
  E_m(X,S) =& \sum_{x \in X}{\alpha s(x)} \nonumber \\ 
  & + \sum_{y \in R_m(S)}{ \beta \Big( { (1/2+ |F_{r}^{(i)}(y)|-c_2) \cdot \sum_{x_j \in O_{r}^{(i)}(y)}{x_j} } + \sum_{ \substack{x_j,x_k \in O_{r}^{(i)}(y) \\ j<k} }{x_jx_k} \Big) } \nonumber \\
  & + \sum_{x \in X}{\gamma g(S,x)}
  \label{eq:energy-family}
\end{align}
\todo{Que vaux $X$ ici ?}	
where $g(S,x)$ denotes a similarity term as distance to initial shape $S$ or data fidelity term and $s(x)$ denotes a length penalization term. Each choice of $m$ generates a different flow, which is generaly described in the Digital Curve Evolution (DCE) algorithm \ref{alg:evolution-model}.

@@

  \label{tab:flow-summary}	  
  \caption{  Since the curvature is negated when reversing the curve (i.e. $\bar{\kappa}=-\kappa$), this process can only shrink  convex parts in shrink mode and expand concave parts in expansion mode.}

@@

\begin{figure}[!ht]
\center
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=3, \alpha=0$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r3-nolength/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=3, \alpha=0.15$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r3-length0.15/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, \alpha=0.15$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r5/summary_flow.eps}
	}%
\end{minipage}%
\caption{The algorithm is very sensitive to the little variations of the estimator, which are particularly important in regions of low squared curvature. Artifacts are somewhat reduced with a length penalization but increases if we use a higher ball radius. }
\label{fig:m1-square-flow}
\end{figure}

@@

\begin{figure}[!ht]
\center
\center
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=1$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l1/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=2$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l2/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=3$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l3/summary_flow.eps}
	}%
\end{minipage}%
\caption{By positioning the estimation ball on farther rings, we minimize artifacts creation. \label{fig:mx-square-flow}}
\end{figure}

In our experiments, the best results are obtained by executing DCE
algorithm with $m$ equal to $r$, where $r$ is the estimation ball
radius (see figures \ref{fig:mx-square-flow},
\ref{fig:mx-flow-gs-radius-effect} ). In plot
\ref{fig:mx-elastica-plots}, we observe how the elastica energy evolves along
the iterations. If the radius is too large, the elastica energy may increase after some steps. That is what happens for the
triangle, in which the flow converges to a single point.

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage[utf8]{inputenc}
\usepackage{graphicx,subfig}
\usepackage{amstext,amsmath,amssymb,bm,bbm,mathtools}
\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
\usepackage[export]{adjustbox}
\usepackage[dvipsnames]{xcolor}

@@

\newcommand{\todo}[1]{{\textcolor{blue}{#1}}}
\newcommand{\test}[1]{{\textcolor{red}{#1}}}
\newcommand{\jaco}[1]{{\textcolor{green!50!black}{#1}}}
\newcommand{\HT}[1]{{\textcolor{violet}{#1}}}
\newcommand{\daniel}[1]{\textcolor{NavyBlue}{#1}}

@@

\begin{document}
%
\title{An Elastica-driven Digital Curve Evolution Model for Image Segmentation\thanks{This  work has  been  partially  funded by CoMeDiC ANR-15-CE40-0006 research grant.}}
  \titlerunning{Digital Curve Evolution by Elastica Energy}
  %\title{Digital Curvature Evolution Models for Image Segmentation\thanks{This  work has  been  partially  funded by CoMeDiC ANR-15-CE40-0006 research grant.}}

\author{Daniel Antunes\inst{1}
Jacques-Olivier Lachaud\inst{1}
Hugues Talbot\inst{2}}
%
\authorrunning{D. Antunes et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Universit{\'e} Savoie Mont Blanc, LAMA, UMR CNRS 5127, F-73376, France
\email{daniel.martins-antunes, jacques-olivier.lachaud@univ-savoie.fr} \and
CentraleSupelec Inria \'equipe OPIS, Universit\'e Paris-Saclay, 9 rue Joliot-Curie, F-91190 Gif-sur-Yvette \\
\email{hugues.talbot@centralesupelec.fr}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  Geometric priors have been shown to be useful in image segmentation to regularize results. For example, the classical
    Mumford-Shah functional uses regions perimeter as prior. This has inspired much research in the last few decades,
    with classical approaches like the Rudin-Osher-Fatemi and most graph-cut formulations, which all use a weighted or
    binary perimeter prior. It has been observed that this prior is not suitable in many applications, for example for
    segmenting thin objects or some textures, which may have high perimeter/surface ratio. Mumford observed that an
    interesting prior for natural objects is the Euler Elastical model, which involves the squared curvature. In other
    areas of science, researchers have noticed that some physical binarization processes, like emulsion unmixing can be
    well-approximated by curvature-related flow like the Willmore flow. 
    However, curvature-related flows are not easy to compute because curvature is difficult to estimate accurately, and the
    underlying optimisation processes are not convex. In this article, we propose to formulate a digital flow that
    approximates a curvature-related flow using a multigrid-convergent curvature estimator, within a discrete
    variational framework. We also present an application of this model as a post-processing step in a segmentation
    framework.
\keywords{Multigrid convergence  \and Digital estimator \and Curvature \and Shape Optimization \and Image Segmentation.}
\end{abstract}
%
%
%
\setcounter{footnote}{0}
\section{Introduction} %% \todo{(To be extended)}}

Geometric quantities are particularly useful as regularizers in
low-level image analysis, especially when little prior information is
known about the shape of interest. Length penalization is a well-behaved, general
purpose regularizer and many models in the literature make use
of it, from active contours \cite{casseles97geodesic} to level-set
formulations \cite{malladi1995image,malladi1995shape}. Discrete
graph-based variational models have been particularly succesfull to
incorporate length penalization, while keeping the
ability to extract a global optimum
\cite{boykov01graphcut,appleton05geodesic}.

However length regularization shows limitations when segmenting
small or thin and elongated objects, as it tends to shrink solutions
or yields disconnected solutions. Such drawbacks can be problematic in
image segmentation or image inpainting. It is thus rather natural to
consider curvature (especially squared curvature) as a potential
regularizer. Energies involving both length and squared curvature are often called
{\em Euler elastica} model (that dates back to Euler) or {\em Willmore energy}
($n$-dimensional extension) and have been brought back to attention in
computer vision by Mumford \cite{mumford1994elastica}. A similar
regularizer is also present in the second-order regularizer
of the original snake model \cite{kass1988snakes}. Similarly,
an explicit curvature term was often employed in early level-set methods,
like~\cite{malladi1995image,malladi1995shape,ballester01filljoint} for segmentation or inpainting.
However, Casselles {\em et al.} in~\cite{casseles97geodesic} observed that these
terms were not geometric, in the sense that they depended on the discretization.

@@

One of the first successful uses of curvature in image processing is the inpainting algorithm
described in \cite{masnou98inpainting}, where authors evaluate the absolute curvature along the level lines of a simply-connected image to reconstruct its occluded parts. The non-intersection property of level lines induces the
construction of an efficient dynamic programming algorithm. However,  the curvature energy is only coarsely
approximated.  In \cite{chan02elasticainpainting}, a geometric inpainting method involving elastica
optimized with level-sets is described.

In image segmentation, Zehiry and Grady have shown that injecting curvature as a regularizer can help recover thin
and elongated objects \cite{zehiry10fast}. Similarly Schoenemann {\em et al.}  \cite{schoenemann09linear} have
considered the elastica energy in their ratio-based image segmentation approach. In~\cite{schoenemann2011elastic},
Schoenemann~{\em et al.} extend ratio-cut approaches to segmentation with a curvature term. Their global optimization
framework shows the considerable advantages of using such regularizer in common binary image segmentation tasks.
However the time complexity of their algorithm is prohibitive, even if curvature is again coarsely
approximated.  In more recent work, Lim {\em et al.} propose an edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set formulation.  They observe improved results compared
to~\cite{casseles97geodesic} and other perimeter-based priors, but observe that a shape prior and a close-enough
solution must also be provided to achieve good results.

In fact, it is still a very challenging task to inject curvature in
the context of image segmentation.  
State-of-art methods are in practice difficult to optimize and do not scale
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In
order to achieve reasonable running times, such approaches make use of
coarse curvature estimations for which the approximation error is
often unknown. Improving the quality of the curvature estimator has an
important impact on the accuracy of the results, but is
computationally too costly in these methods.

Segmentation methods using Elastica or Willmore energies generally
reduce to computing the corresponding flow, i.e. following its
gradient. Well-founded level-set methods involving Willmore energy
\cite{droske2004level} have been used in medical image segmentation
\cite{lim2012introducing}. Another level-set formulation similar to
the Chan-Vese model was proposed in \cite{zhu2013image}. Willmore
flow can also be carried out by phase-field models (see
\cite{bretin2015phase}) but they are less suited to image segmentation
since interfaces are blurred in such models. Threshold dynamics can
also be considered for such energies \cite{esedoglu2008threshold} and
has been proposed for image desocclusion \cite{esedoglu2005threshold}.

In this paper we are interested in purely discrete variational segmentation models involving an elastica
energy. Recently, new estimators of curvature along digital contours have been proposed
\cite{roussillon11mdca,coeurjolly13integral,schindele17mdca} with the desirable multigrid convergence property. This
motivates us to propose models in which they can be used successfully.

In the remainder, we investigate the use of the digital integral invariant
curvature estimator \cite{coeurjolly13integral} in a discrete
variational segmentation framework. More precisely, we show how to
incorporate it in a digital flow minimizing its squared
curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO
algorithm \cite{rother07qpbo}. Our solution does not compute a global
solution to the whole domain, but our approach leads to a digital flow
similar to continuous Elastica, as shown by our experiments. We
present an application of the model as a post-processing step in a
segmentation framework, and demonstrate how it can improve standard
results. \daniel{This paper is an extended version of a previous work that appeared in \cite{antunes19}. We've included a deeper discussion on the optimization method we've used and we present a comparison of our segmentation method with two other related works, and illustrate the results with several examples. Moreover, we have included a new local combinatorial optimization approach for elastica minimization. Finnaly,  the code produced in this paper is freely available on github\footnote{https://www.github.com/danoan/BTools}.}

@@

\section{Multigrid convergent estimators}

The goal of this section is to  introduce the concept of multigrid convergenc and its potential when analyzing digital images with variational models involving geometric quantities. More precisely we wish to delineate shapes within digital images with some priors related to their unknown continuous geometry. Of course we only observe shape digitizations within images.

@@

\begin{figure}[!h]
\center
	\subfloat[\label{}]{%
	\includegraphics[scale=0.2]{images/local_search/definitions/model-regions-square.eps}
	}\hspace{40pt}%
	\subfloat[\label{}]{%
	\includegraphics[scale=0.2]{images/local_search/definitions/glued-curve.eps}
	}%	
	\caption{The blue pixels in figure (a) illustrates a $2$-ring set, while the green pixels its inner-pixel boundary. In figure (b) we highlight an element of set $\mathcal{G}_{11}(C_1,C_2)$. The curve segments of $C_1$ and $C_2$ are colored in red and green, while the junction linels in blue.}
\end{figure}

@@

\begin{figure}[!h]
\center
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h1/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h05/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h1/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h05/summary-flow.eps}
	}%
		\caption{Local combinatorial optimization process results for the square and flower shapes. Shapes displayed at every $2$ iterations for the square and $7$ iterations for the flower.}	
		\label{fig:local-comb-square-results}
\end{figure}

@@

\begin{figure}[!h]
\begin{minipage}[b]{0.5\textwidth}
\center
\includegraphics[scale=0.35]{images/local_search/square-flower.eps}
\caption{Elastica computation along iterations of algorithm \ref{alg:local-search}}
\label{fig:plot-elastica-local-search}
\end{minipage}\hspace{40pt}%
\begin{minipage}[b]{0.45\textwidth}
\center
\captionsetup{type=table}
\begin{tabular}{|l|c|c|}
\hline
& $h=1.0$ & $h=0.5$\\
\hline
Square & 47s (5s/it) & 260s (12s/it)\\
\hline
Flower & 1235s (23s/it)  & 8516s (52s/it)\\
\hline
\end{tabular}
\caption{Running times for the local combinatorial optimization algorithm with $k=1,N=20$. Four threads were used.}
\label{tab:summary-local-comb-rtime}
\end{minipage}
\end{figure}

@@

We define $O_{r}^{(i)}(y) := O^{(i)} \cap B_r(y)$ and an analogous definition holds for $F_{r}^{(i)}(y)$. Expanding \eqref{eq:curvature-estimator-pixels}, we get 
\begin{align*}
  \hat{\kappa}^2(y) &= c_1\Big( c_2 - \sum_{x_j \in B_r(y)} \big( {x_j} + |F_{r}^{(i)}(y)| \big) \Big)^2 \\
   &= c_1 \cdot \Big( C + 2\left( |F_{r}^{(i)}(y)| - c_2 \right) \hspace{-2mm}\sum_{x_j \in O_{r}^{(i)}(y)}\hspace{-2mm}{x_j} + \hspace{-2mm}\sum_{x_j \in O_{r}^{(i)}(y)}\hspace{-2mm}{x_j^2} + \hspace{-2mm}\sum_{ \substack{x_j,x_k \in O_{r}^{(i)}(y) \\ j<k} }\hspace{-2mm}{2x_jx_k}  \Big),
\end{align*}
where $C=c_2^2 - 2c_2 \cdot |F_{r}^{(i)}(y)| + |F_{r}^{(i)}(y)|^2$ is a constant. By ignoring constants and multiplication factors and using
the binary character of the variables, we define the following family
of energies for given parameters $\alpha,\beta, \gamma \geq 0$.
\begin{align}
  E_m(X,S) =& \sum_{x \in X}{\alpha s(x)} \nonumber \\ 
  & + \sum_{y \in R_m(S)}{ \beta \Big( { (1/2+ |F_{r}^{(i)}(y)|-c_2) \cdot \sum_{x_j \in O_{r}^{(i)}(y)}{x_j} } + \sum_{ \substack{x_j,x_k \in O_{r}^{(i)}(y) \\ j<k} }{x_jx_k} \Big) } \nonumber \\
  & + \sum_{x \in X}{\gamma g(S,x)},
  \label{eq:energy-family}
\end{align}
where $g(S,x)$ denotes a similarity term as distance to initial shape $S$ or data fidelity term and $s(x)$ denotes a length penalization term. Each choice of $m$ generates a different flow, which is generaly described in the Digital Curve Evolution (DCE) algorithm \ref{alg:evolution-model}.

@@

  \caption{  Since the curvature is negated when reversing the curve (i.e. $\bar{\kappa}=-\kappa$), this process can only shrink  convex parts in shrink mode and expand concave parts in expansion mode.}
   \label{tab:flow-summary}

@@

\begin{figure}[!h]
\center
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=3, \alpha=0$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r3-nolength/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=3, \alpha=0.15$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r3-length0.15/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, \alpha=0.15$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r5/summary_flow.eps}
	}%
\end{minipage}%
\caption{The algorithm is very sensitive to the little variations of the estimator, which are particularly important in regions of low squared curvature. Artifacts are somewhat reduced with a length penalization but increases if we use a higher ball radius. }
\label{fig:m1-square-flow}
\end{figure}

@@

\begin{figure}[!h]
\center
\center
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=1$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l1/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=2$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l2/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=3$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l3/summary_flow.eps}
	}%
\end{minipage}%
\caption{By positioning the estimation ball on farther rings, we minimize artifacts creation. \label{fig:mx-square-flow}}
\end{figure}
